---
title: Strong expectations cancel locality effects
subtitle: "Code and analysis"
engine: knitr
output: html
toc: true
---

```{r}
#| label: setup
#| output: false

library(car)
library(gdata)
library(ggplot2)
library(lme4)
library(plyr)
library(reshape)
library(knitr)

data <- read.table("data/data.txt", header = F)

colnames(data) <- c("subj",
                    "expt",
                    "item",
                    "cond",
                    "pos",
                    "word",
                    "region",
                    "rt")
```

# Experiment 1
The experiment `RC1` refers to the replication of Levy et al. (2013).
```{r}
## RC1: Levy et al Russian RCs replication
#RC1: Distance manipulaton between relative pronoun and the relative clause verb in Subject RCs and Object RCs. Replication of the Levy Russian RC paper.

#No. of conditions: 4
#Condition a: Subject RC, Long (Canonical word-order)
#Condition b: Subject RC, Short (Non-Canonical word-order)
#Condition c: Object RC, Long (Canonical word-order)
#Condition d: Object RC, Short (Non-Canonical word-order)

#Total items: 24

#Critical region: RCVerb

RC1 <- subset(data, expt == "RC1")

## nested contrasts:
RC1$dist.SR  <- ifelse(RC1$cond == "a",  1,
                ifelse(RC1$cond == "b", -1, 0))
RC1$dist.OR  <- ifelse(RC1$cond == "c",  1,
                ifelse(RC1$cond == "d", -1, 0))
RC1$RCType   <- ifelse(RC1$cond%in%c("a", "b"), -1, 1)

## anova contrasts:
RC1$dist <- ifelse(RC1$cond %in% c("a", "c"),  1, -1)
RC1$int  <- ifelse(RC1$cond %in% c("a", "d"), -1,  1)

questions.RC1      <- subset(RC1, word %in% c("Y", "N"))
questions.RC1$acc  <- as.numeric(as.character(questions.RC1$region))
questions.RC1$cond <- drop.levels(questions.RC1$cond)

mean.accs.RC1 <- round(100*with(questions.RC1, tapply(acc, cond, mean)), digits = 0)
```

## Question response analyses
### Responses accuracy
```{r}
m0.acc.aov <- glmer(
  acc~dist+RCType+int+(1+dist+RCType+int|subj)+(1+dist+RCType+int|item),
  questions.RC1,
  family = binomial()
)

fixefs <- fixef(m0.acc.aov)
SEs    <- sqrt(diag(vcov(m0.acc.aov)))
zs     <- fixefs/SEs

e1acc <- round(cbind(fixefs, SEs, zs),digits = 2)[2:4, 1:3]

rownames(e1acc) <- c("Distance", "RC Type", "DistxRC Type")
colnames(e1acc) <- c("coef.", "SE", "z-score")
colnames(e1acc) <- c("coef.", "SE", "z-score")
```

```{r}
#| echo: false

kable(e1acc, caption = "The main effect of distance, of RC Type, and their interaction on question-response accuracy in Experiment 1.", label = "e1acc2")
```

### Response time
```{r}
m0.resp.RT <- lmer(
  log(rt)~dist.SR+dist.OR+RCType+(1+dist.SR+dist.OR+RCType|subj)+(1+dist.SR+dist.OR+RCType|item),
  questions.RC1
)
```

```{r}
#| echo: false

mean.q.resp <- round(with(questions.RC1, tapply(rt, cond, mean)), digits = 0)

fixefs_qrt <- fixef(m0.resp.RT)
SEs_qrt    <- sqrt(diag(vcov(m0.resp.RT)))
zs_qrt     <- fixefs/SEs

e1qresprt <- round(cbind(fixefs_qrt, SEs_qrt, zs_qrt), digits = 2)[2:4, 1:3]

kable(
  e1qresprt,
  caption = "Effect of distance within subject relatives and object relatives, and main effect of relative clause type on question response time, in Experiment 1."
)
```

## Analysis of critical region
### Frequentist analysis
*ANOVA contrasts:*
```{r}
## crit region RCVerb:
## remove irrelevant tags:
RC1 <- subset(RC1, region!="1" & region!="0" & region!="-")

RC1$region <- drop.levels(RC1$region)

RC1.orig <- RC1 # this isn't really the original RC1 /Johan

## box-cox transform:
RC1$rrt <- -1000/RC1$rt
## we will use log rt:
RC1$lrt <- log(RC1$rt)

m0 <- lmer(
  lrt~dist+RCType+int+(1+dist+RCType+int|subj)+(1|item),
  subset(RC1, region == "RCVerb" & rrt>-10)
)

fixefs <- round(fixef(m0)[2:4], digits = 3)
SEs    <- round(sqrt(diag(vcov(m0)))[2:4], digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt1RCVerbresults <- cbind(fixefs, SEs, ts)
rownames(expt1RCVerbresults) <- c("Distance", "RC Type", "Distance x RC Type")
colnames(expt1RCVerbresults) <- c("coef.", "SE", "t-value")

## we could be discussing predictions in terms of preRCverb, but complicates presentation for reader:
# potentially irrelevant analysis /Johan
m0.pre <- lmer(
    lrt~dist+RCType+int+(1+dist+RCType+int|subj)+(1+dist+RCType+int|item),
    subset(RC1, region == "PreRCVerb" & rrt>-10)
)

fixefs.pre <- round(fixef(m0.pre)[2:4],digits = 3)
SEs.pre <- round(sqrt(diag(vcov(m0.pre)))[2:4],digits = 3)
ts.pre <- round(fixefs.pre/SEs.pre, digits = 2)

expt1RCVerbresults.pre <- cbind(fixefs.pre, SEs.pre, ts.pre)
rownames(expt1RCVerbresults.pre) <- c("Distance", "RC Type", "Distance x RC Type")
colnames(expt1RCVerbresults.pre) <- c("coef.", "SE", "t-value")

precrit.crite1 <- rbind(expt1RCVerbresults, expt1RCVerbresults.pre)
```

```{r}
#| echo: false

kable(
  precrit.crite1,
  caption = "The main effect of distance, of RC Type, and their interaction on reading times at the pre-critical and critical regions in Experiment 1.",
  label = "tab:e1RCVerb"
)

```

*Nested contrasts:*
```{r}
m2.nested <- lmer(
    lrt~dist.SR+dist.OR+RCType+(1+dist.SR+dist.OR+RCType|subj)+(1+dist.SR+dist.OR+RCType|item),
    subset(RC1, region == "RCVerb" & rrt>-10)
)

fixefs <- round(fixef(m2.nested)[2:4],digits = 3)
SEs    <- round(sqrt(diag(vcov(m2.nested)))[2:4],digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt1RCVerbresultsnested <- cbind(fixefs, SEs, ts)
rownames(expt1RCVerbresultsnested) <- c("Distance [SR]", "Distance [OR]", "RC Type")  
colnames(expt1RCVerbresultsnested) <- c("coef.", "SE", "t-value")
```

```{r}
#| echo: false

kable(
  expt1RCVerbresultsnested,
  caption = "The main effect of distance within SRs and ORs, and the main effect of RC Type at the critical region in Experiment 1.",
  label = "tab:e1RCVerbnested"
)
```

## Plots
Figure numbering is consistent with paper.
```{r}
## Data preparation for plotting:
a.regions <- c("NP1", "RCPn", "RCObj",     "PreRCVerb", "RCVerb", "PreMainVerb", "MainVerb")
b.regions <- c("NP1", "RCPn", "PreRCVerb", "RCVerb",    "RCObj",  "PreMainVerb", "MainVerb")
c.regions <- c("NP1", "RCPn", "RCSubj",    "PreRCVerb", "RCVerb", "PreMainVerb", "MainVerb")
d.regions <- c("NP1", "RCPn", "PreRCVerb", "RCVerb",    "RCSubj", "PreMainVerb", "MainVerb")

regions   <- c(a.regions, b.regions, c.regions, d.regions)
region.id <- rep(1:7, 4)
cond.id   <- rep(letters[1:4], each = 7)

region.df <- data.frame(
  cond      = factor(cond.id),
  region.id = region.id,
  region    = factor(regions)
)

#merge multiple "PreRCVerb", "PreMainVerb" regions
##log rt:
RC1.uniq.reg <- ddply(
  RC1, 
  .(subj, expt, item, cond, region), 
  summarize, 
  rt = sum(lrt)
)

RC1.merged <- merge(
  RC1.uniq.reg, region.df, 
  by.x = c("cond", "region")
)

RC1.merged$cond <- drop.levels(RC1.merged$cond)

## define names for factor levels:
RCType <- factor(
  ifelse(
    RC1.merged$cond %in% c("a", "b"),
    "src",
    "orc"
  ),
  levels = c("src", "orc"))

Dist <- factor(
  ifelse(
    RC1.merged$cond %in% c("a", "c"),
    "distant",
    "near"
  ),
  levels = c("near", "distant")
)

RC1.merged$RCType <- RCType
RC1.merged$Dist   <- Dist

## plot log rts:
data.rs <- melt(
  RC1.merged, 
  id = c("RCType", "Dist", "cond", "region", "region.id", "subj"), 
  measure = c("rt"),
  na.rm = TRUE
)

data.id <- data.frame(
  cast(
    data.rs,
    subj+RCType+Dist+cond+region+region.id ~ .,
    function(x) c(rt = mean(x), N = length(x))
  )
)

## compute means and corrected CIs:
GM <- mean(tapply(data.id$rt, data.id$subj, mean))

data.id <- ddply(data.id, .(subj), 
                 transform, rt.w = rt - mean(rt) + GM)  

temp <- melt(data.id, id.var = c("subj", "RCType", "Dist", "cond", "region", "region.id"), measure.var = "rt.w")

M.id.w <- cast(
  temp,
  RCType+Dist+cond+region+region.id ~ ., 
  function(x) c(M = mean(x), SE = sd(x)/sqrt(length(x)), N = length(x) ) )
```

```{r}
#| label: fig-expt1-rcverb
#| fig-cap: "Experiment 1: Reading times in log ms at the critical region (relative clause verb) for the four conditions."
#| warning: false

pd <- position_dodge(width = 0.05)
k <- 1
                                        
(p1a <- ggplot(subset(M.id.w, region == "RCVerb"), 
               aes(x = Dist, y = M, group = RCType)) +
   geom_point(shape = 21, fill = "white", size = k*3) +
   geom_line(aes(linetype = RCType), size = k) +
   geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                 width = .1, size = k) +
   xlab("Distance") +
   ylab("reading time [log ms]") +
   scale_colour_hue(name = "relative clause type", # Legend label, use darker colors
                    breaks = c("src", "orc"),
                    labels = c("SRC", "ORC"),
                    l = 40)+                     
   labs(title = "Critical region [RC verb]") +
   theme_bw() +
   labs(legend.position = c(.87, .6))
)
```

``` {r}
M.id.w.src <- subset(M.id.w, RCType == "src")
M.id.w.src$RCType <- drop.levels(M.id.w.src$RCType)

M.id.w.orc <- subset(M.id.w, RCType == "orc")
M.id.w.orc$RCType <- drop.levels(M.id.w.orc$RCType)

## function for generic by-region plots:
byregion.plot.sr <- function(
    data,
    mytitle,
    k = 1,
    x.lab = "position",
    y.lab = "reading time [log ms]"
) {
  ggplot(data,
         aes(x = region.id, y = M, group = Dist)) + 
     geom_point(shape = 21, size = k*2) +
     geom_line(aes(linetype = Dist),
               size = k) +
     geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                   width = .1,
                   size = k) +
     #xlab(x.lab)+
     scale_x_continuous(x.lab, breaks = 1:7, labels = c("NP1", "RCPn", "PreRCVerb\nRCObj", "RCVerb\nPreRCVerb", "RCObj\nRCVerb", "PreMainVerb", "MainVerb")) +
     ylab(y.lab) +             
     ggtitle(mytitle) +
     theme_bw() +
     theme(axis.text.x = element_text(size = 7))
}

byregion.plot.or <- function(
    data,
    mytitle,
    k = 1,
    x.lab = "position",
    y.lab = "reading time [log ms]"
  ) {
  ggplot(data,
         aes(x = region.id, y = M, group = Dist)) + 
     geom_point(shape = 21,
                size = k*2) +
     geom_line(aes(linetype = Dist),
               size = k) +
     geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                   width = .1,
                   size = k)+
     #xlab(x.lab)+
     scale_x_continuous(x.lab, breaks = 1:7, labels = c("NP1", "RCPn", "PreRCVerb\nRCSubj", "RCVerb\nPreRCVerb", "RCSubj\nRCVerb", "PreMainVerb", "MainVerb"))+
     ylab(y.lab)+             
     ggtitle(mytitle) +
     theme_bw()+
     theme(axis.text.x = element_text(size = 7))
}
```

```{r}
#| label: fig-expt1-disteff-sr
#| fig-cap: "Experiment 1: Reading times by region in subject relatives. The critical region is RCVerb."
#| warning: false

## SRC plot:
(plot.src <- byregion.plot.sr(
  M.id.w.src,
  mytitle = "Effect of distance [subject relatives]",
  k = 1,
  x.lab = "position",
  y.lab = "reading time [log ms]")
)
```

```{r}
#| label: fig-expt1-disteff-or
#| fig-cap: "Experiment 1: Reading times by region in object relatives. The critical region is RCVerb."
#| warning: false

(plot.orc <- byregion.plot.or(
  M.id.w.orc,
  mytitle = "Effect of distance [object relatives]", k = 1,
  x.lab = "position", y.lab = "reading time [log ms]")
)
```

```{r}
## generic function for RC type comparison:
byregion.plot.rc.distant <- function(
    data,
    mytitle,
    k = 1,
    x.lab = "position",
    y.lab = "reading time [log ms]"
  ) {
  ggplot(data,
         aes(x = region.id, y = M, group = RCType)) + 
    geom_point(shape = 21,
               size = k*2) +
    geom_line(aes(linetype = RCType),
              size = k) +
    geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                  width = .1,
                  size = k) +
    #xlab(x.lab)+
    scale_x_continuous(x.lab, breaks = 1:7, labels = c("NP1", "RCPn", "RCObj\nRCSubj", "PreRCVerb", "RCVerb", "PreMainVerb", "MainVerb"))+
    ylab(y.lab)+             
    ggtitle(mytitle) +
    theme_bw()+
    theme(axis.text.x = element_text(size = 7))
}

byregion.plot.rc.near <- function(
    data,
    mytitle,
    k = 1,
    x.lab = "position",
    y.lab = "reading time [msec]"){
  ggplot(data,
         aes(x = region.id, y = M, group = RCType)) + 
    geom_point(shape = 21,
               size = k*2) +
    geom_line(aes(linetype = RCType),
              size = k) +
    geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                  width = .1,
                  size = k) +
    #xlab(x.lab)+
    scale_x_continuous(x.lab, breaks = 1:7, labels = c("NP1", "RCPn", "PreRCVerb", "RCVerb", "RCObj\nRCSubj", "PreMainVerb", "MainVerb"))+
    ylab(y.lab)+             
    ggtitle(mytitle) +
    theme_bw()+
    theme(axis.text.x = element_text(size = 7))
}
```

```{r}
#| layout-ncol: 2

## compare SR vs OR:
M.id.w.dist <- subset(M.id.w, Dist == "distant")
M.id.w.dist$Dist <- drop.levels(M.id.w.dist$Dist)

M.id.w.near <- subset(M.id.w, Dist == "near")
M.id.w.near$Dist <- drop.levels(M.id.w.near$Dist)

# these two plots are not in the paper /Johan
(plot.rctype.dist <- byregion.plot.rc.distant(
  M.id.w.dist,
  mytitle = "Effect of RC type (distant conditions)",
  k = 1,
  x.lab = "position", y.lab = "reading time [msec]")
)


(plot.rctype.near <- byregion.plot.rc.near(
  M.id.w.near,
  mytitle = "Effect of RC type (near conditions)",
  k = 1,
  x.lab = "position", y.lab = "reading time [msec]")
)

## the diff at the premain verb is due to diff. words
premain <- subset(RC1, region == "PreMainVerb" & rrt>-10 & cond %in% c("a", "c"))
```
These two plots were not included in the paper.

# Experiment 2
```{r}
CP1 <- subset(data, expt == "CP1")

a.regions <- c("PreNounPred", "CPNounPred", "PreLV", "CPLightVerb", "PreCoordMV", "CoordMainVerb")
b.regions <- c("PreNounPred", "CPNounPred", "PreLV", "CPLightVerb", "PreCoordMV", "CoordMainVerb")
c.regions <- c("PreMVObj",    "MVObj",      "PreMV", "MainVerb",    "PreCoordMV", "CoordMainVerb")
d.regions <- c("PreMVObj",    "MVObj",      "PreMV", "MainVerb",    "PreCoordMV", "CoordMainVerb")

# set up data frame for further analysis
regions <- c(a.regions, b.regions, c.regions, d.regions)
region.id <- rep(1:6, 4)
cond.id   <- rep(letters[1:4], each = 6)

region.df <- data.frame(
  cond      = factor(cond.id),
  region.id = region.id,
  region    = factor(regions)
)

## nested contrasts:
CP1$dist.exp   <- ifelse(CP1$cond == "a", -1, ifelse(CP1$cond == "b", 1, 0))
CP1$dist.noexp <- ifelse(CP1$cond == "c", -1, ifelse(CP1$cond == "d", 1, 0))

## anova contrasts:
CP1$dist <- ifelse(CP1$cond %in% c("a", "c"), -1,  1)
CP1$exp  <- ifelse(CP1$cond %in% c("a", "b"),  1, -1)
CP1$int  <- ifelse(CP1$cond %in% c("a", "d"), -1,  1)

CP1$lrt <- log(CP1$rt)
```

## Sentence completion study analysis
```{r}
completiondata <- read.table("data/sentence_completion_stats.txt", header = TRUE)
## the correct verbs:
verbs <- completiondata$verb
## For each row in df, look up the verb, then 
## choose the column that corresponds to that 
## verb and save the proportion:
n <- dim(completiondata)[1]
pr <- rep(NA,n)
for(i in 1:n){
  vb <- completiondata[i,]$verb
  pr[i] <- completiondata[i, vb]
}


probabilities <- data.frame(
  item  = rep(1:16, each = 4),
  cond  = rep(letters[1:4],16),
  probs = rep(pr, each = 2)
)

meanprobs <- with(probabilities, tapply(probs, cond, mean))

## noun frequencies:
npfreq <- read.table("data/expt2-freq.txt", header = T)
with(npfreq, tapply(nounfreq, condition, mean))
with(npfreq, tapply(nounfreq, condition, sd))/sqrt(16)

conda <- subset(npfreq, condition == "a")
condc <- subset(npfreq, condition == "c")

t.test(conda$nounfreq, condc$nounfreq, paired = T)
```

## Question response analyses
### Response accuracy
```{r}
#Experiment 4 (Complex predicate)

## question response accuracies:
questions.CP1 <- subset(CP1, word %in% c("Y", "N"))
questions.CP1$acc  <- as.numeric(as.character(questions.CP1$region))
questions.CP1$cond <- drop.levels(questions.CP1$cond)
mean.accs.CP1      <- round(100*with(questions.CP1, tapply(acc, cond, mean)), digits = 0)

# fit accuracy model with ANOVA contrasts /Johan
m0.acc.aov <- glmer(acc~dist+exp+int+(1+dist+exp+int|subj)+(1+dist+exp+int|item), questions.CP1, family = binomial())

fixefs <- fixef(m0.acc.aov)
SEs    <- sqrt(diag(vcov(m0.acc.aov)))
zs     <- fixefs/SEs

e2acc <- round(cbind(fixefs, SEs, zs), digits = 2)[2:4, 1:3]

rownames(e2acc) <- c("Distance", "RC Type", "DistxRC Type")
colnames(e2acc) <- c("coef.", "SE", "z-score")
```

```{r}
#| echo: false

kable(
  e2acc,
  caption = "The main effect of distance, of RC Type, and their interaction on question-response accuracy in Experiment 2."
)
```

### Response time
```{r}
m0.rt.aov <- lmer(log(rt)~dist+exp+int+(1+dist+exp+int|subj)+(1|item), questions.CP1) # this response time model is never used /Johan

# added the same table as everywhere else for completeness /Johan
fixefs <- fixef(m0.rt.aov)
SEs    <- sqrt(diag(vcov(m0.rt.aov)))
zs     <- fixefs/SEs

e2qresprt <- round(cbind(fixefs, SEs, zs), digits = 2)[2:4, 1:3]

rownames(e2qresprt) <- c("Distance", "RC Type", "DistxRC Type")
colnames(e2qresprt) <- c("coef.", "SE", "z-score")
```

```{r}
#| echo: false

kable(
  e2qresprt,
  caption = "The main effect of distance, of RC Type, and their interaction on question-response accuracy in Experiment 2."
)
```


## Plots
```{r}
## Data preparation for plotting:
CP1 <- subset(CP1, region!="1" & region!="0" & region!="-")

CP1$region <- drop.levels(CP1$region)
CP1$cond   <- drop.levels(CP1$cond)

#update labels in the long conditions (PreLV1, PreMV1)
CP1$region[CP1$region == 'PreMV1'] <- 'PreMV'
CP1$region[CP1$region == 'PreLV1'] <- 'PreLV'

#Labeling error in the data; item == 15, cond == a, b. 'PreMV' should be 'PreLV'
CP1$region[CP1$region == 'PreMV' & CP1$item == 15 & CP1$cond == 'a'] <- 'PreLV'
CP1$region[CP1$region == 'PreMV' & CP1$item == 15 & CP1$cond == 'b'] <- 'PreLV'

## crit region: CPLightVerb/MainVerb:
#merge multiple "PreNounPred", "PreMVObj", "PreLV", "PreMV", "PreCoordMV" regions
CP1.uniq.reg <- ddply(
  CP1,
  .(subj, expt, item, cond, region, dist, exp, int, dist.exp, dist.noexp),
  summarize,
  rt = sum(rt)
)

CP1.uniq.reg.merged <- merge(
  CP1.uniq.reg, region.df,
  by.x = c("cond", "region")
)

CP1.uniq.reg.merged$cond <- drop.levels(CP1.uniq.reg.merged$cond)

#CP1: Expectation build-up for the verb due to nominal host of Complex Predicate. Effect of surprisal/locality on Normal vs Complex predicate.

#No. of conditions: 4
#Condition a: Expectation, Long
#Condition b: Expectation, Short
#Condition c: No Expectation, Long
#Condition d: No Expectation, Short

#Total items: 16

CP1.uniq.reg.merged$Exp  <- factor(ifelse(CP1.uniq.reg.merged$cond %in% c("a", "b"), "exp", "no-exp"), levels = c("exp", "no-exp"))
CP1.uniq.reg.merged$Dist <- factor(ifelse(CP1.uniq.reg.merged$cond %in% c("b", "d"), "short", "long"), levels = c("short", "long"))

### log rt:
CP1.uniq.reg.merged$lrt <- log(CP1.uniq.reg.merged$rt)
data.rs <- melt(
  CP1.uniq.reg.merged,
  id = c("Exp", "Dist", "cond", "region", "region.id", "subj"), 
  measure = c("lrt"),
  na.rm = TRUE
)

CP1 <- CP1.uniq.reg.merged

#Critical region: CPLightVerb
## these are all region.id == 4:
CP1.ab.crit <- subset(CP1, region == "CPLightVerb")
CP1.cd.crit <- subset(CP1, region == "MainVerb")
CP1.crit    <- rbind(CP1.ab.crit, CP1.cd.crit)

CP1.crit$region2 <- factor("crit")
```

```{r}
CP1.postcrit <- subset(CP1, region.id == 5)
CP1.postcrit$region2 <- factor("postcrit")

CP1.postcrit2 <- subset(CP1, region.id == 6)
CP1.postcrit2$region2 <- factor("postcrit2")

CP1.critdata <- rbind(CP1.crit, CP1.postcrit, CP1.postcrit2)
CP1.critdata <- merge(CP1.critdata, probabilities, by.x = c("item", "cond"), by.y = c("item", "cond"))

pred <- ifelse(CP1.critdata$probs < median(CP1.critdata$probs), "low", "high")

CP1.critdata$pred <- factor(pred)

data.id <- data.frame(
  cast(
    data.rs,
    subj+Exp+Dist+cond+region+region.id ~ ., 
    function(x) c(lrt = mean(x), N = length(x))
  )
)                     

GM <- mean(tapply(data.id$lrt, data.id$subj, mean))

#removing between subject variance
data.id <- ddply(
  data.id,
  .(subj),
  transform,
  lrt.w = lrt - mean(lrt) + GM
)  

temp <- melt(data.id, id.var = c("subj", "Exp", "Dist", "cond", "region", "region.id"), measure.var = "lrt.w")

M.id.w <- cast(
  temp,
  Exp+Dist+cond+region+region.id ~ .,
  function(x) c(M = mean(x), SE = sd(x)/sqrt(length(x)), N = length(x))
) 

pd <- position_dodge(width = 0.05)
k <- 1

#rename 'CPLightVerb' in condition (a, b) in order to draw the plot
M.id.w$region[M.id.w$region == 'CPLightVerb'] <- 'MainVerb'

MVregion <- subset(M.id.w, region == "MainVerb")
```

```{r}
#| label: fig-expt2-crit
#| fig-cap: "Experiment 2: Reading times in log ms at the critical region (light/simple verb) for the four conditions."

(p1a <- ggplot(subset(M.id.w, region == "MainVerb"), 
               aes(x = Dist, y = M, group = Exp)) + 
    geom_point(shape = 21, fill = "white", size = k*3) +
    geom_line(aes(linetype = Exp),size = k) +
    geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                  width = .1,
                  size = k) +
    xlab("Distance") +
    ylab("reading time [log ms]") +
    scale_colour_hue(name = "Expectation condition", # Legend label, use darker colors
                     breaks = c("exp", "no-exp"),
                     labels = c("EXP", "NO-EXP"),
                     l = 40) +                     
    ggtitle("Critical region [Finite Verb]") +
    theme_bw() 
)
```

```{r}
## subset by Expectation condition:
M.id.w.exp <- subset(M.id.w, Exp == "exp")
M.id.w.exp$Exp <- drop.levels(M.id.w.exp$Exp)

M.id.w.noexp <- subset(M.id.w, Exp == "no-exp")
M.id.w.noexp$Exp <- drop.levels(M.id.w.noexp$Exp)

## function for generic by-region plots:
byregion.plot.cp <- function(
    data,
    mytitle,
    k = 1,
    x.lab = "position",
    y.lab = "reading time [msec]"
) {
  ggplot(data,
         aes(x = region.id, y = M, group = Dist)) + 
  geom_point(shape = 21, size = k*2) +
  geom_line(aes(linetype = Dist), size = k) +
  geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                width = .1,
                size = k) +
  scale_x_continuous(x.lab, breaks = 1:6, labels = c("PreNounPred", "CPNounPred", "PreLV", "CPLightVerb", "PreCoordMV", "CoordMainVerb")) +
  ylab(y.lab) +             
  ggtitle(mytitle) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 7))
}

byregion.plot.sp <- function(
    data,
    mytitle, k = 1,
    x.lab = "position",
    y.lab = "reading time [msec]")
{
  ggplot(data,
         aes(x = region.id, y = M, group = Dist)) + 
  geom_point(shape = 21, size = k*2) +
  geom_line(aes(linetype = Dist), size = k) +
  geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                width = .1,
                size = k) +
  #xlab(x.lab)+
  scale_x_continuous(x.lab, breaks = 1:6, labels = c("PreMVObj", "MVObj", "PreMV", "MainVerb", "PreCoordMV", "CoordMainVerb")) +
  ylab(y.lab) +             
  ggtitle(mytitle) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 7))
}
```

```{r}
#| label: fig-expt2-disteff-cp
#| fig-cap: "Experiment 2: Reading times by region in the complex predicate condition. The critical region is CPLightVerb."

(plot.exp <- byregion.plot.cp(
  M.id.w.exp,
  mytitle = "Effect of distance [Complex predicate condition]",
  k = 1,
  x.lab = "position",
  y.lab = "reading time [log ms]"
))
```

```{r}
#| label: fig-expt2-disteff-sp
#| fig-cap: "Experiment 2: Reading times by region in the simple predicate condition. The critical region is MainVerb."

(plot.noexp <- byregion.plot.sp(
  M.id.w.noexp,
  mytitle = "Effect of distance [Simple predicate condition]", k = 1,
  x.lab = "position",
  y.lab = "reading time [log ms]"
))
```

```{r}
## subset by Distance condition:
M.id.w.short <- subset(M.id.w, Dist == "short")
M.id.w.short$Dist <- drop.levels(M.id.w.short$Dist)

M.id.w.long <- subset(M.id.w, Dist == "long")
M.id.w.long$Dist <- drop.levels(M.id.w.long$Dist)

## function for generic by-region plots:
byregion.plot2.short <- function(
    data,
    mytitle,
    k = 1,
    x.lab = "position",
    y.lab = "reading time [msec]"
) {
  ggplot(data, aes(x = region.id, y = M, group = Exp)) + 
  geom_point(shape = 21,
             size = k*2) +
  geom_line(aes(linetype = Exp),
            size = k) +
  geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                width = .1,
                size = k) +
                  #xlab(x.lab)+
  scale_x_continuous(x.lab, breaks = 1:6, labels = c("PreNounPred\nPreMVObj", "CPNounPred\nMVObj", "PreLV\nPreMV", "CPLightVerb\nMainVerb", "PreCoordMV", "CoordMainVerb"))+
  ylab(y.lab)+             
  ggtitle(mytitle) +
  theme_bw()+
  theme(axis.text.x = element_text(size = 7))
}

byregion.plot2.long <- function(
    data,
    mytitle, k = 1,
    x.lab = "position",
    y.lab = "reading time [msec]"
) {
  ggplot(data,
         aes(x = region.id, y = M, group = Exp)) + 
  geom_point(shape = 21,
             size = k*2) +
  geom_line(aes(linetype = Exp),
            size = k) +
  geom_errorbar(aes(ymin = M-2*SE, ymax = M+2*SE),
                width = .1,
                size = k) +
                  #xlab(x.lab)+
  scale_x_continuous(x.lab, breaks = 1:6, labels = c("PreNounPred\nPreMVObj", "CPNounPred\nMVObj", "PreLV\nPreMV", "CPLightVerb\nMainVerb", "PreCoordMV", "CoordMainVerb"))+
  ylab(y.lab)+             
  ggtitle(mytitle) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 7))
}
```
not used in paper; should be removed?

## Summary for Johan paper: SHOULD BE (RE-)MOVED
```{r}
head(CP1.crit)
means <- with(CP1.crit, tapply(rt, cond, mean))
means
dist <- rep(c("long", "short"), 2)
expectation <- rep(c("expected", "unexpected"), each = 2)
means2 <- data.frame(dist, expectation, means)
means2
```

## Reading time analyses
### Frequentist 
#### Critical region
##### ANOVA contrast
Using factorial coding for high and low expectation:
```{r}
m01.anova.crit.asfactor <- lmer(lrt~dist+exp+int+(1|subj)+(1|item), subset(CP1.critdata, region2 == "crit"))

fixefs <- round(fixef(m01.anova.crit.asfactor)[2:4],digits = 3)
SEs    <- round(sqrt(diag(vcov(m01.anova.crit.asfactor)))[2:4],digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt4aovcrit <- cbind(fixefs, SEs, ts)

rownames(expt4aovcrit) <- c("Distance", "Expectation", "Dist x Exp")
colnames(expt4aovcrit) <- c("coef.", "SE", "t-value")

kable(expt4aovcrit,
       caption = "The main effect of distance, of expectation, and their interaction on reading times at the critical region in Experiment 2.",
       label = "tab:e4critaov")
```

Using raw probabilities instead of factorial coding:
```{r}
## highly correlated:
# xyplot(probs~exp|item, CP1.critdata) # xyplot is not included in any of the imported packages /Johan

m01.anova.crit.asprob <- lmer(
  lrt~dist*scale(probs, scale = F)+(1|subj)+(0+scale(probs, scale = F)|subj)+(0+dist:scale(probs, scale = F)|subj)+(1|item),
  subset(CP1.critdata, region2 == "crit")
)

fixefs <- round(fixef(m01.anova.crit.asprob)[2:4], digits = 3)
SEs    <- round(sqrt(diag(vcov(m01.anova.crit.asprob)))[2:4], digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt4aovcrit <- cbind(fixefs, SEs, ts)
rownames(expt4aovcrit) <- c("Distance", "Expectation", "Dist x Exp")
colnames(expt4aovcrit) <- c("coef.", "SE", "t-value")

kable(expt4aovcrit,
       caption = "The main effect of distance, of expectation (using probabilities), and their interaction on reading times at the critical region in Experiment 2.",
       label = "tab:e4critaov2")
```

##### Nested contrast
```{r}
m0.nested.crit <- lmer(
  lrt~(dist.exp+dist.noexp+exp)*scale(probs, scale = F)+(1+(dist.exp+dist.noexp+exp)*scale(probs, scale = F)|subj)+(1+dist.exp+dist.noexp+exp|item),
  subset(CP1.critdata, region2 == "crit")
)

fixefs <- round(fixef(m0.nested.crit)[2:4], digits = 3)
SEs    <- round(sqrt(diag(vcov(m0.nested.crit)))[2:4], digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt4nestedcrit <- cbind(fixefs, SEs, ts)
rownames(expt4nestedcrit) <- c("Distance (Exp)", "Distance (No Exp)", "Exp")
colnames(expt4nestedcrit) <- c("coef.", "SE", "t-value")

kable(expt4aovcrit,
       caption = "Effect of distance")
```

#### Post-critical region
*ANOVA contrasts:*
```{r}
# which one of these was actually of interest? /Johan
m0.anova.postcrit <- lmer(
  lrt~dist+exp+int+(1|subj)+(1|item),
  subset(CP1.critdata, region2 == "postcrit")
)

m0.anova.postcrit <- lmer(
  lrt~dist+exp+int+(dist+exp+int-1|subj)+(1|subj)+(1|item),
  subset(CP1.critdata, region2 == "postcrit")
)  

m0.anova.postcrit <- lmer(
  lrt~dist*scale(probs, scale = F)+(1+dist*scale(probs, scale = F)|subj)+(1+dist*scale(probs, scale = F)|item),
  subset(CP1.critdata, region2 == "postcrit")
)  

fixefs <- round(fixef(m0.anova.postcrit)[2:4], digits = 3)
SEs    <- round(sqrt(diag(vcov(m0.anova.postcrit)))[2:4], digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt4aovpostcrit <- cbind(fixefs, SEs, ts)
rownames(expt4aovpostcrit) <- c("Distance", "Expectation", "Dist x Exp")
colnames(expt4aovpostcrit) <- c("coef.", "SE", "t-value")
```

*Nested contrasts:*
```{r}
m0.nested.postcrit <- lmer(
  lrt~(dist.exp+dist.noexp+exp)*scale(probs, scale = F)+(1+(dist.exp+dist.noexp+exp)*scale(probs, scale = F)|subj)+(1+dist.exp+dist.noexp+exp|item),
  subset(CP1.critdata, region2 == "postcrit")
)

fixefs <- round(fixef(m0.nested.postcrit)[2:4], digits = 3)
SEs    <- round(sqrt(diag(vcov(m0.nested.postcrit)))[2:4], digits = 3)
ts     <- round(fixefs/SEs, digits = 2)

expt4nestedpostcrit <- cbind(fixefs, SEs, ts)
rownames(expt4nestedpostcrit) <- c("Distance (Exp)", "Distance (No Exp)", "Exp")
colnames(expt4nestedpostcrit) <- c("coef.", "SE", "t-value")
```

#### Region 2 (CPNounPred/MVObj)
*ANOVA contrasts:*
```{r}
CP1.pos2 <- subset(CP1, region.id == 2)

## anova contrasts:
CP1.pos2$dist <- ifelse(CP1.pos2$cond %in% c("a", "c"), -1,  1)
CP1.pos2$int  <- ifelse(CP1.pos2$cond %in% c("a", "d"), -1,  1)
CP1.pos2$exp  <- ifelse(CP1.pos2$cond %in% c("a", "b"),  1, -1)

CP1.pos2$lrt <- log(CP1.pos2$rt)

m0 <- lmer(lrt~exp+dist+int+(1+exp+dist+int|subj)+(1+exp+dist+int|item), CP1.pos2)
```
